---
title: "Baseline groups stress"
author: "Paul A. Bloom"
date: "2023-07-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Setup

```{r}
library(Matrix)
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
library(brms)
library(cowplot)
library(bayestestR)
library(EMAtools)
```


```{r}
load('../cleaned_data/cleaned_data_for_stress_social_context_project_2024-12-16.rda')


# recode to have -1=SI/SA, 0=Control for contrasts
ema_data = mutate(ema_data, Control=-ifelse(GROUP=='Control', 0, 1))

# keep control=1, SI/SA=1 for selfreport data (strain)
selfreport = mutate(selfreport, Control = ifelse(GROUP == 'Control', 0, 1))

```


# 1. Case-control differences in social context: fit multinomial logistic regression for 3-level categorical outcome (peers, fam, alone)

```{r, multinom_case_control_model, message=FALSE}
# student t prior with 3df, center at 0, scale parameter 10
beta_prior_logit = c(prior_string("student_t(3, 0, 10)", class = "b"))

multinom_alone_case_control = brms::brm(data = ema_data, timeWith_cat3 ~ Control + SITE + sex + age + 
                           weekend + timebin +  schoolclosure + personal_time + (1|ID), 
                           family = categorical(link = 'logit'),
                           prior = beta_prior_logit, 
                           cores = 4, seed = 11291993, iter = 4000)
```

## Case-control differences in affect, stress

```{r, affect_stress_models, message=FALSE}
# student t prior with 3df, center at 0, scale parameter 100 (for non-scaled EMA outcomes on a 1-100 scale)
ema_beta_100_prior = c(prior_string("student_t(3, 0, 100)", class = "b"))

stress_case_control = brms::brm(data = ema_data, stressed ~ Control + weekend + personal_time + timebin + 
                                summerbreak + SITE + sex + age + (1|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000, 
                              prior = ema_beta_100_prior)

na_case_control = brms::brm(data = ema_data, mean_na ~ Control + weekend + personal_time + timebin + 
                                summerbreak + SITE + sex + age + (1|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                            prior = ema_beta_100_prior)

pa_case_control = brms::brm(data = ema_data, mean_pa ~ Control + weekend + personal_time + timebin + 
                                summerbreak + SITE + sex + age + (1|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                            prior = ema_beta_100_prior)


```

### Print model outputs
```{r}
#stres
fixef(stress_case_control)

#na
fixef(na_case_control)

#pa
fixef(pa_case_control)


# cohen's d type effect sizes
# https://imaging.mrc-cbu.cam.ac.uk/statswiki/FAQ/tdunpaired
sigma_stressed = sqrt(mean(c(var(ema_data$stressed[ema_data$Control==0], na.rm = TRUE),
                           var(ema_data$stressed[ema_data$Control==-1], na.rm = TRUE))))

sigma_na = sqrt(mean(c(var(ema_data$mean_na[ema_data$Control==0], na.rm = TRUE),
                     var(ema_data$mean_na[ema_data$Control==-1], na.rm = TRUE))))


sigma_pa = sqrt(mean(c(var(ema_data$mean_pa[ema_data$Control==0], na.rm = TRUE),
                     var(ema_data$mean_pa[ema_data$Control==-1], na = TRUE))))


# stress ES
fixef(stress_case_control)/sigma_stressed

# na ES
fixef(na_case_control)/sigma_na

# pa ES
fixef(pa_case_control)/sigma_pa

```

# School closure interaction analyses with group on stress
```{r, schoolclosuremodels, message=FALSE}
ema_data = mutate(ema_data, school_closure_3 = case_when(
  schoolclosure == 1 ~ 'during',
  schoolclosure == 0 & date < as.Date('2020-06-04') ~ 'before',
  schoolclosure == 0 & date > as.Date('2021-01-01') ~ 'after',
))


# school closure as binary variable
stress_case_control_inter_schoolclosure1 = brms::brm(data = ema_data %>% mutate(Control = as.character(Control)), stressed ~ Control*schoolclosure + weekend + personal_time + timebin + 
                                summerbreak + SITE + sex + age + (schoolclosure|ID), cores = 4, seed = 11291993, iter = 4000, 
                              prior = ema_beta_100_prior)

# school closure as before/during/after
stress_case_control_inter_schoolclosure = brms::brm(data = ema_data %>% mutate(Control = as.character(Control)), stressed ~ Control*school_closure_3 + weekend + personal_time + timebin + 
                                summerbreak + SITE + sex + age + (school_closure_3|ID), cores = 4, seed = 11291993, iter = 4000, 
                              prior = ema_beta_100_prior)

```

Summarize and plot school closure models

```{r}
summary(stress_case_control_inter_schoolclosure1)
summary(stress_case_control_inter_schoolclosure)

sc3_interaction_condfx = conditional_effects(stress_case_control_inter_schoolclosure)
covid_stress_inter_plot = sc3_interaction_condfx[[10]] %>%
  mutate(group = ifelse(Control == -1, 'STB', 'Psychiatric Control')) %>%
  ggplot(data = ., aes(x = school_closure_3, y = estimate__, color = group)) +
  geom_point(size = 4, position = position_dodge(0.2)) +
  geom_errorbar(aes(ymin = lower__, ymax = upper__, width = 0), position = position_dodge(0.2), lwd = 1) +
  theme_bw() +
  labs(x = 'Timing Relative to Covid-19 School Closure', y = 'Estimated Mean Stress', color = NULL) +
  scale_x_discrete(limits = c('before', 'during', 'after'))


ggsave(covid_stress_inter_plot, file = '../figures/supplement/covid_stress_interaction.png', height = 4, width=6)
```


## Case-control differences in prior adversity exposure

```{r}
strain_characteristics = c('CIAllCT', 'CIAllTH', 'CDAllCT', 'CDAllTH', 'CHAllCT', 'CHAllTH', 'CEAllCT', 'CEAllTH', 'CRAllCT', 'CRAllTH',
                           'StressCT', 'StressTH', 'EvntCT', 'DiffCT', 'EvntTH','DiffTH')

selfreport = mutate(selfreport,
                       # Count
                       RecDAllCT = RecDDiffCT + RecEvntCT,
                       RecEAllCT = RecEDiffCT + RecEvntCT,
                       RecHAllCT = RecHDiffCT + RecEvntCT,
                       RecLAllCT = RecLDiffCT + RecEvntCT,
                       RecRAllCT = RecRDiffCT + RecEvntCT,
                       # Severity
                       RecDAllTH = RecDDiffTH + RecEvntTH,
                       RecEAllTH = RecEDiffTH + RecEvntTH,
                       RecHAllTH = RecHDiffTH + RecEvntTH,
                       RecLAllTH = RecLDiffTH + RecEvntTH,
                       RecRAllTH = RecRDiffTH + RecEvntTH
                       )

strain_characteristics_pastyear = c('RecDAllCT', 'RecEAllCT', 'RecHAllCT', 'RecLAllCT', 'RecRAllCT',
                                    'RecDAllTH', 'RecEAllTH', 'RecHAllTH', 'RecLAllTH', 'RecRAllTH', 
                                    'RecTotCT', 'RecTotTH', 'RecEvntCT', 'RecEvntTH', 'RecDiffCT', 'RecDiffTH')


case_control_strain_models = selfreport %>%
    pivot_longer(all_of(c(strain_characteristics, strain_characteristics_pastyear))) %>%
    group_by(name) %>%
    mutate(value_scale = scale(value)) %>%
    nest() %>%
    mutate(strain_model = purrr::map(data, ~lm(data = ., 
                                                       value_scale ~ Control + SITE + sex + age)),
           strain_coef = purrr::map(strain_model, ~broom.mixed::tidy(., conf.int=TRUE)))

case_control_strain_coef = case_control_strain_models %>%
  unnest(strain_coef) %>%
  dplyr::select(-data, -strain_model) %>%
  dplyr::filter(term == 'Control') %>%
  dplyr::mutate(characteristic = case_when(
        startsWith(name, 'CI') ~ 'Interpersonal Loss',
        startsWith(name, 'CD') ~ 'Physical Danger',
        startsWith(name, 'CH') ~ 'Humiliation',
        startsWith(name, 'CE') ~ 'Entrapment',
        startsWith(name, 'CR') ~ 'Role Change/Reversal',
        startsWith(name, 'Stress') ~ 'Total Stressors',
        startsWith(name, 'Evnt') ~ 'Total Acute Life Events',
        startsWith(name, 'Diff') ~ 'Total Chronic Difficulties',
        name == 'RecDAllTH' ~ 'Physical Danger',
        name == 'RecEAllTH' ~ 'Entrapment',
        name == 'RecHAllTH'~ 'Humiliation', 
        name == 'RecLAllTH' ~ 'Interpersonal Loss',
        name =='RecRAllTH' ~ 'Role Change/Reversal',
        name =='RecTotTH' ~ 'Total Stressors',
        name =='RecEvntTH' ~ 'Total Acute Life Events',
        name == 'RecDiffTH'~ 'Total Chronic Difficulties'
    ),
    strain_timing = ifelse(startsWith(name, 'Rec'), 'Past Year', 'Lifetime'),
    strain_type = case_when(
        endsWith(name, 'TH') ~ 'Severity',
         endsWith(name, 'CT') ~ 'Count'),
    strain_total = case_when(
      grepl('Total', characteristic) ~ 'Total Summaries',
      !grepl('Total', characteristic) ~ 'Core Social-Psych\nCharacteristics'
    ))
```

## Save all models
```{r}
save(multinom_alone_case_control, stress_case_control, na_case_control, pa_case_control, 
     case_control_strain_models, case_control_strain_coef,
     file = '../model_outputs/case_control_models.rda')

```

# 2. Plot case-control differences!
```{r}
load('../model_outputs/case_control_models.rda')
```

## Social Context Plot / Stats from Multinomial Logistic Regression Model
```{r}
summary(multinom_alone_case_control)

# fixef
fixef(multinom_alone_case_control)
```

```{r}
# prediction grid for social context
pred_grid_case_control = expand.grid(Control = -1:0, 
                                     weekend = 0, 
                                     personal_time = mean(ema_data$personal_time, na.rm = TRUE), 
                                     summerbreak = 0,
                                     schoolclosure = 0,
                          SITE = 'CUIMC', sex= 'F', age = mean(ema_data$age, na.rm=TRUE), timebin = '1morning')

```

Stats for social context contrasts
```{r}
# convert probability to odds
probability_to_odds = function(probability){
  return(probability / (1-probability))
}

# using the prediction grid, get fitted predictions for each condition, then manually calculate contrasts
multinom_cond_fx_draws = fitted(multinom_alone_case_control, newdata = pred_grid_case_control, 
                                re_formula = NA, incl_autocor=TRUE, summary = FALSE, scale = 'response', categorical=TRUE) %>%
  as.data.frame() %>%
  mutate(diff_alone = `2.alone` - `1.alone`,
         diff_family = `2.family` - `1.family`,
         diff_peer = `2.peer` - `1.peer`,
         or_alone = probability_to_odds(`2.alone`) / probability_to_odds(`1.alone`),
         or_family = probability_to_odds(`2.family`) / probability_to_odds(`1.family`),
         or_peer = probability_to_odds(`2.peer`) / probability_to_odds(`1.peer`)) 

# 95% HDI for contrasts
multinom_cond_fx_draws %>%
  bayestestR::hdi(ci=0.95) %>%
  as.data.frame()

# means for contrasts
multinom_cond_fx_draws %>%
  colMeans() %>%
  as.data.frame() 
```


```{r}
# extract conditional (marginal) probabilities for PC vs. SI/SA
multinom_cond_fx = conditional_effects(multinom_alone_case_control, categorical=TRUE, scale='response', robust = FALSE, 
                                       re_formula = NA, method = 'fitted')

multinom_predictions = multinom_cond_fx$`Control:cats__` %>%
    dplyr::filter(Control %in% c(0,-1)) %>%
  mutate(group = dplyr::recode(Control, '0'='Psychiatric Controls', '-1'='STB'))

# participant-level summaries of raw data (for spaghetti)
timewith_raw_summary = ema_data %>%
  dplyr::filter(!is.na(Control)) %>%
  group_by(ID, Control, PCTalone, PCTpeer, PCTfamily) %>%
  count() %>%
  pivot_longer(contains('PCT')) %>%
  mutate(`cats__`=gsub('PCT', '', name)) %>%
  dplyr::select(ID, Control, `cats__`, `estimate__`=value) %>%
  mutate(group = dplyr::recode(Control, '0'='Psychiatric Controls', '-1'='STB'))


# plot
case_control_alone_plt = ggplot(multinom_predictions, aes(x = `cats__`, y = `estimate__`, color = group)) +
    geom_line(data = timewith_raw_summary, aes(group = ID),  alpha = 0.1) +
    geom_point(position = position_dodge(0.1), size = 3) +
    geom_errorbar(aes(ymin = `lower__`, ymax = `upper__`), width = 0.1, position = position_dodge(0.1), lwd = 1) +
    geom_line(aes(group = group), position = position_dodge(0.1), lwd = 1) +
    labs(y = '% of Responses', x = 'Social Context in Last 2h', color = 'Group',
         title = 'Case-Control Differences: Social Context') +
    theme_bw()

``` 

## Stress & Affect Plot

```{r}
pred_grid_case_control = expand.grid(Control = 0:1, weekend = 0, personal_time = 0, summerbreak = 0,
                          SITE = 'PITT', sex= 'M', age = 15, timebin = '3late_afternoon')

pred_grid_case_control_stress = fitted(stress_case_control, newdata = pred_grid_case_control, re_formula = NA, incl_autocor=FALSE) %>%
  data.frame() %>%
  cbind(pred_grid_case_control, .)


case_control_stress_affect_plt = rbind(broom.mixed::tidy(stress_case_control) %>% mutate(outcome = 'Stress'),
      broom.mixed::tidy(na_case_control) %>% mutate(outcome = 'Negative Affect'),
      broom.mixed::tidy(pa_case_control) %>% mutate(outcome = 'Positive Affect')) %>%
  dplyr::filter(term == 'Control') %>%
  ggplot(data = ., aes(x = outcome, y = estimate)) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0, lty = 2) + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0, lwd = 1) +
  labs(y = 'Mean Difference (0-100 Scale)\nSTB+ > Psychiatric Controls', x = NULL,
       title = 'Case-Control Differences\nMomentary Stress & Affect') +
  theme_bw() +
  coord_flip()
      
```


## Strain variables plot
```{r}
case_control_strain_plot = case_control_strain_coef %>% dplyr::filter(strain_type == 'Severity') %>%
  ggplot(data = ., aes(x = characteristic, y = estimate, color = strain_timing)) +
  geom_point(position = position_dodge(0.3), size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0, 
                position = position_dodge(0.3),
                lwd = 1) +
  theme_bw() +
  coord_flip() +
  facet_grid(rows = vars(strain_total), scales = 'free_y') +
  geom_hline(yintercept = 0, lty = 2) +
  labs(color = 'Metric', x = NULL, y = 'Standardized Mean Difference\nSTB+ > Psychiatric Controls',
       title = 'Case-Control Differences:\nPrior Stress & Adversity Exposure Severity') +
  scale_color_manual(values = c('black', 'red4'))
```


## Example Data Plot
```{r}
example_data_plot = dplyr::filter(ema_data, ID=='1101', Day<=7) %>%
  group_by(Day) %>%
  mutate(min_x = ifelse((Timedif > 120 | is.na(Timedif)), hour - 2, hour - (Timedif/60))) %>%
  ungroup() %>%
  pivot_longer(c('mean_pa', 'mean_na', 'stressed')) %>%
  mutate(day = dplyr::recode(Day, '1'='Mon', '2'='Sat', '3'='Sun', '4'='Thurs', '5'='Fri', '6'='Tues', '7'='Wed'),
         name = dplyr::recode(name, 'stressed'='Stress', 'mean_pa'='Positive\nAffect', 'mean_na'='Negative\nAffect'),
         day =  factor(day, levels = c('Mon', 'Tues', 'Wed', 'Thurs', 'Fri', 'Sat', 'Sun'))) %>%
  group_by(name) %>%
  mutate(min_y = min(value),
         max_y = max(value)) %>%
  ungroup() %>%
  ggplot(aes(x = hour, y = value)) +
  facet_grid(rows = vars(name), cols = vars(day), scales = 'free_y', switch = 'both') +
  geom_rect(aes(fill = timeWith_cat3, ymin = min_y, ymax = max_y, xmin = min_x, xmax = hour), alpha = 0.5) +
  geom_point() + 
  geom_line() +
  theme_bw() +
  theme(panel.spacing = unit(0, 'cm'),
        panel.grid.minor = element_blank(),
        legend.position = 'top') +
  labs(x = 'Hour (24 Hour Clock)', y = NULL, title = 'Example Experience Sampling Data', fill = 'Social Context in Last 2h') +
  scale_fill_viridis_d() 


```

## Put together grid of plots

```{r}
case_control_plot = cowplot::plot_grid(example_data_plot, case_control_alone_plt,
                   labels = c('A', 'B'))


cowplot::save_plot(case_control_plot, filename = '../figures/main/1_case_control_plot.png', base_width = 10, base_height = 4.25)
```


# Differences in individual ema items

```{r}
# separate models for each item
nested = ema_data %>%
  pivot_longer(c('angry', 'anxious', 'calm', 'confident', 'happy', 'included', 'lonely', 'rejected', 'sad', 'stressed', 'supported')) %>%
  group_by(name, ID) %>%
  ungroup() %>%
  group_by(name) %>%
  nest() %>%
  mutate(model = purrr::map(data, ~lme4::lmer(data =., value ~ Control + weekend + personal_time + timebin + 
                                summerbreak + SITE + sex + age + (1|ID))),
         estimates = purrr::map(model, ~emmeans::emmeans(., specs = 'Control') %>%
                                  as.data.frame()),
         stats = purrr::map(model, ~emmeans::emmeans(., specs = 'Control') %>%
                              pairs(adjust=FALSE) %>%
                                  as.data.frame()))

unnnested = nested %>%
  unnest(estimates) %>%
  dplyr::select(-model, -data, -stats) %>%
  mutate(Control = ifelse(Control =='-1', 'STB', 'Psychiatric Control'))

unnnested_stats = nested %>%
  unnest(stats) %>%
  dplyr::select(-model, -data, -estimates) %>%
  mutate(., sig = p.value<.05)


unnnested = left_join(unnnested, unnnested_stats %>% dplyr::select(name, sig), by = 'name') %>%
  mutate(sig = ifelse(sig==TRUE & Control == 'STB', '*', ''))

single_item_case_control = ggplot(unnnested, aes(x = name, y = emmean)) +
  geom_point(position = position_dodge(0.35), size = 3, aes(color = Control)) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL, color = Control), width = 0, position = position_dodge(0.35), lwd =1) +
  theme_bw() + 
  labs(y = 'Estimated Group Means', x = 'Individual EMA Items', color = 'Group') +
  theme(text = element_text(face='bold')) +
  scale_color_manual(values = c('#F8766D', '#619CFF')) +
  geom_text(aes(label=sig), size = 14)

ggsave(single_item_case_control, file = '../figures/supplement/single_item_case_control.png', height = 4, width = 9)


```
