---
title: "Social Context, Affect, and Stress"
author: "Paul A. Bloom"
date: "2023-07-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Time spent alone vs with others {.tabset}

## Setup

```{r}
library(Matrix)
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
library(brms)
load('../cleaned_data/cleaned_data_for_stress_social_context_project_2024-12-16.rda')

# student t prior with 3df, center at 0, scale parameter 10
beta_prior_logit = c(prior_string("student_t(3, 0, 10)", class = "b"))

# student t prior with 3df, center at 0, scale parameter 100 (for non-scaled EMA outcomes on a 1-100 scale)
ema_beta_100_prior = c(prior_string("student_t(3, 0, 100)", class = "b"))
```


# 1. Affect & Stress ~ Concurrent Social Context (Whole Group)

```{r}
ema_data = mutate(ema_data, 
                  alone_vs_family = case_when(
                    timeWith_cat3 == 'alone' ~ 1,
                    timeWith_cat3 == 'peer' ~ NA_real_,
                    timeWith_cat3 == 'family' ~ 0),
                  alone_vs_peer = case_when(
                    timeWith_cat3 == 'alone' ~ 1,
                    timeWith_cat3 == 'family' ~ NA_real_,
                    timeWith_cat3 == 'peer' ~ 0))

ema_data = mutate(ema_data,
                  weekend = as.numeric(as.character(weekend)),
                  schoolclosure = as.numeric(as.character(schoolclosure)),
                  summerbreak = as.numeric(as.character(summerbreak)),
                  personal_time_z = scale(personal_time))


ema_data_check = ema_data %>%
  group_by(ID) %>%
  summarise(n = n(),
            age = age[1],
            sex = sex[1],
            SITE = SITE[1],
            GROUP=GROUP[1])


schoolclosure_na = ema_data %>% dplyr::filter(is.na(schoolclosure))


ema_data = dplyr::left_join(ema_data, dplyr::select(selfreport, ID, participant_suicide_event1), by = 'ID')

sum(is.na(ema_data$participant_suicide_event1))
```


```{r, echo=FALSE}
stress_social_context = brms::brm(data = ema_data, stressed_within_center ~ timeWith_cat3 + weekend + personal_time_z + schoolclosure + (timeWith_cat3|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

na_social_context = brms::brm(data = ema_data, na_within ~ timeWith_cat3 + weekend + personal_time_z + schoolclosure + (timeWith_cat3|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

pa_social_context = brms::brm(data = ema_data, pa_within ~ timeWith_cat3 + weekend + personal_time_z + schoolclosure + (timeWith_cat3|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)
```

## 1. Affect & Stress ~ Concurrent Social Context (Alone vs. with others)

```{r}
ema_data = ema_data %>%
  group_by(ID) %>%
  mutate(alone_centered = scale(as.numeric(as.character(alone))),
         alone_mean = mean(as.numeric(as.character(alone)), na.rm=T)) %>%
  ungroup()


# preliminary model
m1 = lmerTest::lmer(data = ema_data, stressed ~ alone  +  weekend + personal_time_z  + schoolclosure + SITE + 
                  (alone|ID))

summary(m1)
```

```{r}
stress_social_context_case_control = brms::brm(data = ema_data, stressed ~ alone*Control + weekend + personal_time_z + schoolclosure + SITE + (alone|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

stress_social_context_case_control_no_rslope = brms::brm(data = ema_data, stressed ~ alone*Control + weekend + personal_time_z + schoolclosure + SITE + (1|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

na_social_context_case_control = brms::brm(data = ema_data, mean_na ~ alone*Control + weekend + personal_time_z + schoolclosure + SITE + (alone|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

pa_social_context_case_control = brms::brm(data = ema_data, mean_pa ~ alone*Control + weekend + personal_time_z + schoolclosure + SITE + (alone|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

stress_social_context_case_control_detail = brms::brm(data = ema_data, stressed ~ timeWith_cat3*Control + weekend + personal_time_z + schoolclosure + SITE + (timeWith_cat3|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

stress_social_context_case_control_detail_no_rslope = brms::brm(data = ema_data, stressed ~ timeWith_cat3*Control + weekend + personal_time_z + schoolclosure + SITE + (1|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

na_social_context_case_control_detail = brms::brm(data = ema_data, mean_na ~ timeWith_cat3*Control + weekend + personal_time_z + schoolclosure + SITE + (timeWith_cat3|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

pa_social_context_case_control_detail = brms::brm(data = ema_data, mean_pa ~ timeWith_cat3*Control + weekend + personal_time_z + schoolclosure + SITE + (timeWith_cat3|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

fixef(pa_social_context_case_control)
fixef(na_social_context_case_control)
fixef(stress_social_context_case_control)
```

# Save model outputs

```{r}
save(stress_social_context, na_social_context, pa_social_context,
     stress_social_context_case_control, na_social_context_case_control, pa_social_context_case_control, 
     stress_social_context_case_control_detail, na_social_context_case_control_detail, pa_social_context_case_control_detail,
     stress_social_context_case_control_no_rslope, stress_social_context_case_control_detail_no_rslope,
     file = '../model_outputs/social_context_stress_affect.rda')
```


```{r}
load('../model_outputs/social_context_stress_affect.rda')

summary(stress_social_context_case_control_no_rslope)
summary(stress_social_context_case_control)
summary(stress_social_context_case_control_detail)


comparisons = 3
corected_quantiles = c(0.025 / comparisons, 1 - (.025 / comparisons))


summary(stress_social_context_case_control)
fixef(object = stress_social_context_case_control, probs = corected_quantiles)

```

# Effect sizes
```{r}
sigma_stressed = sqrt(mean(c(var(ema_data$stressed[ema_data$timeWith_cat3=='peer'], na.rm = TRUE),
                           var(ema_data$stressed[ema_data$timeWith_cat3=='family'], na.rm = TRUE),
                           var(ema_data$stressed[ema_data$timeWith_cat3=='alone'], na.rm = TRUE))))
stress_effectsize_df = emmeans::emmeans(stress_social_context, specs = 'timeWith_cat3') %>%
  pairs(reverse=T) %>%
  as.data.frame() 

stress_effectsize_df = mutate(stress_effectsize_df, d = estimate / sigma_stressed)
stress_effectsize_df
```

```{r}
sigma_na = sqrt(mean(c(var(ema_data$mean_na[ema_data$timeWith_cat3=='peer'], na.rm = TRUE),
                           var(ema_data$mean_na[ema_data$timeWith_cat3=='family'], na.rm = TRUE),
                           var(ema_data$mean_na[ema_data$timeWith_cat3=='alone'], na.rm = TRUE))))
na_effectsize_df = emmeans::emmeans(na_social_context, specs = 'timeWith_cat3') %>%
  pairs(reverse=T) %>%
  as.data.frame() 

na_effectsize_df = mutate(na_effectsize_df, d = estimate / sigma_na)
na_effectsize_df


sigma_pa = sqrt(mean(c(var(ema_data$mean_pa[ema_data$timeWith_cat3=='peer'], na.rm = TRUE),
                           var(ema_data$mean_pa[ema_data$timeWith_cat3=='family'], na.rm = TRUE),
                           var(ema_data$mean_pa[ema_data$timeWith_cat3=='alone'], na.rm = TRUE))))
pa_effectsize_df = emmeans::emmeans(pa_social_context, specs = 'timeWith_cat3') %>%
  pairs(reverse=T) %>%
  as.data.frame() 

pa_effectsize_df = mutate(pa_effectsize_df, d = estimate / sigma_pa)
pa_effectsize_df
```


```{r}
emmeans::emmeans(na_social_context, specs = 'timeWith_cat3') %>%
  pairs(reverse=T)


emmeans::emmeans(pa_social_context, specs = 'timeWith_cat3') %>%
  pairs(reverse=T)

emmeans::emmeans(stress_social_context, specs = 'timeWith_cat3') %>%
  pairs(reverse=T)
```


## 1a: plots for main fx
```{r}
pred_grid_social_context = expand.grid(timeWith_cat3 = c('family', 'peer', 'alone'),
                                       weekend = 0,
                                       schoolclosure = 0,
                                       personal_time_z = 0, 
                                       SITE = 'CUIMC')


pred_grid_social_context_stress = fitted(stress_social_context, newdata = pred_grid_social_context, re_formula = NA, incl_autocor=FALSE) %>%
  data.frame() %>%
  cbind(pred_grid_social_context, .)

pred_grid_social_context_na = fitted(na_social_context, newdata = pred_grid_social_context, re_formula = NA, incl_autocor=FALSE) %>%
  data.frame() %>%
  cbind(pred_grid_social_context, .)

pred_grid_social_context_pa = fitted(pa_social_context, newdata = pred_grid_social_context, re_formula = NA, incl_autocor=FALSE) %>%
  data.frame() %>%
  cbind(pred_grid_social_context, .)


fixef(stress_social_context)
fixef(na_social_context)
fixef(pa_social_context)


social_context_affect_preds = rbind(pred_grid_social_context_stress %>% mutate(outcome = ' Stress'),
      pred_grid_social_context_na %>% mutate(outcome = 'Negative\nAffect'),
      pred_grid_social_context_pa %>% mutate(outcome = 'Positive\nAffect')) %>%
  mutate(timeWith_cat3 = str_to_title(timeWith_cat3))

social_context_affect_plot = ggplot(social_context_affect_preds, aes(x = timeWith_cat3, y = Estimate, color = timeWith_cat3)) +
  geom_hline(yintercept = 0, lty =2, alpha = 0.5) +
  geom_line(aes(group= outcome), color = 'black') +
  geom_errorbar(aes(ymin = Q2.5, ymax = Q97.5), width = 0.1, lwd = 1) +
  geom_point(size = 3) +
  facet_grid(~outcome) +
  theme_bw() +
  theme(legend.position = 'none') +
  scale_color_viridis_d() +
  scale_x_discrete(limits = c('Family', 'Alone', 'Peer')) +
  labs(y = 'Difference From\nParticipant Mean',
       x = NULL,
       title= '') 
    
```
## 2a: plots for case-control interactions

```{r}
pred_grid_social_context_case_control = expand.grid(timeWith_cat3 = c('family', 'peer', 'alone'),
                                       Control = 0:1,
                                       schoolclosure = 0,
                                       weekend = 0,
                                       personal_time_z = 0, 
                                       SITE = 'CUIMC')

fixef(stress_social_context_case_control_detail)
pred_grid_social_context_case_control_stress = fitted(stress_social_context_case_control_detail,
                                                      newdata = pred_grid_social_context_case_control, re_formula = NA, 
                                                      incl_autocor=FALSE) %>%
  data.frame() %>%
  cbind(pred_grid_social_context_case_control, .)

pred_grid_social_context_case_control_na = fitted(na_social_context_case_control_detail, newdata = pred_grid_social_context_case_control, re_formula = NA, incl_autocor=FALSE) %>%
  data.frame() %>%
  cbind(pred_grid_social_context_case_control, .)

pred_grid_social_context_case_control_pa = fitted(pa_social_context_case_control_detail, newdata = pred_grid_social_context_case_control, re_formula = NA, incl_autocor=FALSE) %>%
  data.frame() %>%
  cbind(pred_grid_social_context_case_control, .)


fixef(stress_social_context_case_control)
fixef(na_social_context_case_control)
fixef(pa_social_context_case_control)


social_context_case_control_affect_preds = rbind(pred_grid_social_context_case_control_stress %>% mutate(outcome = ' Stress'),
      pred_grid_social_context_case_control_na %>% mutate(outcome = 'Negative\nAffect'),
      pred_grid_social_context_case_control_pa %>% mutate(outcome = 'Positive\nAffect')) %>%
  mutate(group = case_when(Control == 0 ~ 'STB', 
                           Control ==1 ~ 'Psychiatric Control')) %>%
  mutate(timeWith_cat3 = str_to_title(timeWith_cat3))



case_control_buffering_plot = ggplot(social_context_case_control_affect_preds, aes(x = timeWith_cat3, y = Estimate, color = group)) +
  geom_line(aes(group=group), position = position_dodge(0.2), lwd = 1) +
  geom_errorbar(aes(ymin = Q2.5, ymax = Q97.5), width = 0.1, lwd = 1, position = position_dodge(0.2)) +
  geom_point(size = 3, position = position_dodge(0.2)) +
  facet_grid(~outcome) +
  theme_bw() +
  labs(y = 'Estimated Average\n + 95% CI',
       x = NULL,
       color = NULL, title = '') +
  scale_x_discrete(limits = c('Family', 'Alone', 'Peer')) +
  theme(legend.position = 'right', legend.direction = 'vertical')
      
```




# Differences among participants with events during study


```{r}
stress_social_context_stb_during_study = brms::brm(data = ema_data, stressed ~ alone*participant_suicide_event1 + weekend + personal_time_z + schoolclosure + SITE + (alone|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

stress_social_context_stb_during_study_lmer = lmerTest::lmer(data = ema_data, stressed ~ timeWith_cat3*Control + weekend + personal_time_z + schoolclosure + SITE + (1|ID))
summary(stress_social_context_stb_during_study_lmer)


na_social_context_stb_during_study = brms::brm(data = ema_data, mean_na ~ alone*participant_suicide_event1 + weekend + personal_time_z + schoolclosure + SITE + (alone|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

pa_social_context_stb_during_study = brms::brm(data = ema_data, mean_pa ~ alone*participant_suicide_event1 + weekend + personal_time_z + schoolclosure + SITE + (alone|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

stress_social_context_stb_during_study_detail = brms::brm(data = ema_data, stressed ~ timeWith_cat3*participant_suicide_event1 + weekend + personal_time_z + schoolclosure + SITE + (timeWith_cat3|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

na_social_context_stb_during_study_detail = brms::brm(data = ema_data, mean_na ~ timeWith_cat3*participant_suicide_event1 + weekend + personal_time_z + schoolclosure + SITE + (timeWith_cat3|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

pa_social_context_stb_during_study_detail = brms::brm(data = ema_data, mean_pa ~ timeWith_cat3*participant_suicide_event1 + weekend + personal_time_z + schoolclosure + SITE + (timeWith_cat3|ID) +
                                ar(time = datetime, gr=id_timepoint), 
                              cores = 4, seed = 11291993, iter = 4000,
                              prior = ema_beta_100_prior)

save(stress_social_context_stb_during_study, stress_social_context_stb_during_study_detail,
     na_social_context_stb_during_study, na_social_context_stb_during_study_detail,
     pa_social_context_stb_during_study, pa_social_context_stb_during_study_detail,
     file = '../model_outputs/social_context_stress_affect_participants_stb_during_study.rda')
```


```{r}
load('../model_outputs/social_context_stress_affect_participants_stb_during_study.rda')


fixef(na_social_context_stb_during_study)
round(fixef(na_social_context_stb_during_study_detail),2)

fixef(pa_social_context_stb_during_study)
fixef(pa_social_context_stb_during_study_detail)

fixef(stress_social_context_stb_during_study)
fixef(stress_social_context_stb_during_study_detail)


```

## plot for stb during study vs. no stb
```{r}
pred_grid_social_context_stb_during_study = expand.grid(timeWith_cat3 = c('family', 'peer', 'alone'),
                                       participant_suicide_event1 = 0:1,
                                       schoolclosure = 0,
                                       weekend = 0,
                                       personal_time_z = 0, 
                                       SITE = 'CUIMC')

pred_grid_social_context_stb_during_study_stress = fitted(stress_social_context_stb_during_study_detail,
                                                      newdata = pred_grid_social_context_stb_during_study, re_formula = NA, 
                                                      incl_autocor=FALSE) %>%
  data.frame() %>%
  cbind(pred_grid_social_context_stb_during_study, .)

pred_grid_social_context_stb_during_study_na = fitted(na_social_context_stb_during_study_detail, newdata = pred_grid_social_context_stb_during_study, re_formula = NA, incl_autocor=FALSE) %>%
  data.frame() %>%
  cbind(pred_grid_social_context_stb_during_study, .)

pred_grid_social_context_stb_during_study_pa = fitted(pa_social_context_stb_during_study_detail, newdata = pred_grid_social_context_stb_during_study, re_formula = NA, incl_autocor=FALSE) %>%
  data.frame() %>%
  cbind(pred_grid_social_context_stb_during_study, .)


social_context_stb_during_study_affect_preds = rbind(pred_grid_social_context_stb_during_study_stress %>% mutate(outcome = ' Stress'),
      pred_grid_social_context_stb_during_study_na %>% mutate(outcome = 'Negative\nAffect'),
      pred_grid_social_context_stb_during_study_pa %>% mutate(outcome = 'Positive\nAffect')) %>%
  mutate(group = case_when(participant_suicide_event1 == 1 ~ '1+ Suicidal Events',
                           participant_suicide_event1 == 0 ~ 'No Suicidal Events',)) %>%
  mutate(timeWith_cat3 = str_to_title(timeWith_cat3))



stb_during_study_buffering_plot = ggplot(social_context_stb_during_study_affect_preds, aes(x = timeWith_cat3, y = Estimate, color = group)) +
  geom_line(aes(group=group), position = position_dodge(0.2), lwd = 1) +
  geom_errorbar(aes(ymin = Q2.5, ymax = Q97.5), width = 0, lwd = 1, position = position_dodge(0.2)) +
  geom_point(size = 3, position = position_dodge(0.2)) +
  facet_grid(~outcome) +
  theme_bw() +
  labs(y = 'Estimated Average\n+ 95% CI',
       x = 'Social Context (Last 2h)',
       title = '',
       color = 'Participants with vs. without suicide events\nduring 6-month study window') +
  scale_x_discrete(limits = c('Family', 'Alone', 'Peer')) +
  scale_color_viridis_d(begin = 0.1, end = 0.5) +
  theme(legend.position = 'right', legend.direction = 'vertical')
      

ggsave(stb_during_study_buffering_plot, file = '../figures/draft/stb_during_study_buffering.png', height = 5, width = 8)
```


# Make plot grid
```{r}
fig2_grid = cowplot::plot_grid(social_context_affect_plot, case_control_buffering_plot, stb_during_study_buffering_plot,
                   labels = c('A', 'B', 'C'),
                   label_x = 0, label_y = 1,
                   nrow = 3, align = 'v', axis = 'lr')



cowplot::save_plot(fig2_grid, base_width = 9, base_height = 10, filename = '../figures/main/2_social_context_affect_stress.png')
```

# Individual items
```{r}
nested = ema_data %>%
  pivot_longer(c('angry', 'anxious', 'calm', 'confident', 'happy', 'included', 'lonely', 'rejected', 'sad', 'stressed', 'supported')) %>%
  group_by(name, ID) %>%
  mutate(value_scale = scale(value, center =TRUE, scale=FALSE)) %>%
  ungroup() %>%
  group_by(name) %>%
  nest() %>%
  mutate(model = purrr::map(data, ~lme4::lmer(data =., value_scale ~ timeWith_cat3 + weekend + personal_time_z + schoolclosure + (timeWith_cat3|ID))),
         estimates = purrr::map(model, ~emmeans::emmeans(., specs = 'timeWith_cat3') %>%
                                  as.data.frame()))


unnnested = nested %>%
  unnest(estimates) %>%
  dplyr::select(-model, -data)

single_item_social_context = ggplot(unnnested, aes(x = name, y = emmean, color = timeWith_cat3)) +
  geom_hline(yintercept = 0, color = 'black', lty = 2) +
  geom_point(position = position_dodge(0.35), size = 3) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0, position = position_dodge(0.35), lwd =1) +
  theme_bw() + 
  labs(y = 'Mean Difference From Participant Mean', x = 'Individual EMA Items', color = 'Social Context') +
  scale_color_viridis_d() +
  theme(text = element_text(face='bold'))

ggsave(single_item_social_context, file = '../figures/supplement/single_item_social_context.png', height = 4, width = 9)

```
