---
title: "5_social_buffering_stb"
author: "Paul A. Bloom"
date: "2023-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
library(Matrix)
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
library(metafor)
library(brms)
library(survival)
library(ranger)
library(ggfortify)
library(pscl)
load('../cleaned_data/cleaned_data_for_stress_social_context_project_2024-12-16.rda')

```

```{r}
load('/Volumes/columbia/MAPS_Data/Analysis/Wrangling/STB_Risk/suicide_events_2024-10-23.rda')


selfreport = mutate(selfreport, Control = GROUP == 'Control')


# update events based on Karla's updates fall 2024
suicide_events_clean = dplyr::mutate(suicide_events_clean, start_date = if_else(start_date == '#####' &
                                                                                         event == 'attempt' & 
                                                                                         ptid == '#####', 
                                                                                       as.Date('#####'), as.Date(start_date)),
                                     end_date = if_else(end_date == '#####' &
                                                                                         event == 'attempt' & 
                                                                                         ptid == '#####', 
                                                                                       as.Date('#####'), as.Date(end_date)),
                                     start_date_week_wednesday = if_else(start_date_week_wednesday == '#####' &
                                                                                         event == 'attempt' & 
                                                                                         ptid == '#####', 
                                                                                       as.Date('#####'), as.Date(start_date_week_wednesday)))


suicide_events_clean = dplyr::mutate(suicide_events_clean, start_date = if_else(start_date == '#####' &
                                                                                         event == 'attempt' & 
                                                                                         ptid == '#####', 
                                                                                       as.Date('#####'), as.Date(start_date)),
                                     end_date = if_else(end_date == '#####' &
                                                                                         event == 'attempt' & 
                                                                                         ptid == '#####', 
                                                                                       as.Date('#####'), as.Date(end_date)))

suicide_events_info_report = dplyr::filter(suicide_events_clean, ptid %in% ema_data$ID)

sum_suicide_events_participant = suicide_events_clean %>% group_by(ptid) %>% count()


# unique participants
length(unique(suicide_events_info_report$ptid))

# total events
nrow(suicide_events_info_report)

# table 
table(suicide_events_info_report$event)

# unique weeks
suicide_events_info_weekly_report = dplyr::filter(suicide_events_clean_weekly, ptid %in% ema_data$ID)

# 33 unique weeks (4 instances )
nrow(suicide_events_info_weekly_report)


```


# Run level-1 GLMS to get a "social buffering" metric for each participant
```{r}
ema_data_for_buffer = dplyr::filter(ema_data, Nalone > 0, Npeer > 0, Nfamily > 0, Nresp >= 5) %>%
    group_by(ID) %>%
    mutate(stress_variance = var(stressed)) %>%
    ungroup() %>%
    dplyr::filter(stress_variance > 0)


length(unique(ema_data_for_buffer$ID))
length(unique(ema_data$ID))


nested_buffer_models = ema_data_for_buffer %>%
    group_by(ID) %>%
    mutate(stress_within_z = scale(stressed, scale = TRUE, center = TRUE),
           n=n()) %>%
    group_by(ID, n) %>%
    nest() %>%
    dplyr::mutate(stress_buffer_level1_model = map(data, ~lm(data=., stress_within_z ~ alone + hour + personal_time)))


nested_buffer_coef = nested_buffer_models %>%
    dplyr::mutate(coef = map(stress_buffer_level1_model, ~broom::tidy(.)))


stress_buffer_scores = nested_buffer_coef %>%
    unnest(coef) %>%
    dplyr::select(-stress_buffer_level1_model, -data) %>%
    dplyr::filter(term == 'alone1') %>%
    dplyr::select(ID, stress_buffer_est = estimate, stress_buffer_sd = std.error, n) %>%
    mutate(stress_buffer_se = stress_buffer_sd/sqrt(n))


stress_buffer_selfreport = left_join(selfreport, stress_buffer_scores, by = 'ID')


length(unique(stress_buffer_scores$ID))

```

# Does stress buffering predict which participants will have suicide events?
```{r}
stress_buffer_glm_level2 = glm(data = stress_buffer_selfreport, participant_suicide_event1 ~ stress_buffer_est + GROUP + sex + SITE + age + iPhone, weights = 1/stress_buffer_sd,
                               family = binomial(link = 'logit'))

stress_buffer_glm_level2_si = glm(data = stress_buffer_selfreport, participant_suicide_event1 ~ stress_buffer_est + SSI21_i + sex + SITE + age + iPhone, weights = 1/stress_buffer_sd,
                               family = binomial(link = 'logit'))


summary(stress_buffer_glm_level2)
summary(stress_buffer_glm_level2_si)

```


# Does stress predict which participants will have suicide events?
```{r}
btw_summary_data = ema_data %>%
  mutate(alone = as.numeric(as.character(alone)),
         Wpeers = as.numeric(as.character(Wpeers))) %>%
  group_by(ID) %>%
  summarise(
         baseline_mean_stress = mean(stressed[personal_time < 8], na.rm=TRUE),
         mean_stress = mean(stressed, na.rm=TRUE),
         baseline_mean_na = mean(mean_na[personal_time < 8], na.rm = TRUE),
         baseline_mean_pa = mean(mean_pa[personal_time < 8], na.rm = TRUE),
         mean_na = mean(mean_na, na.rm = TRUE),
         mean_pa = mean(mean_pa, na.rm = TRUE),
         mean_stress_alone = mean(stressed[alone ==1], na.rm=TRUE),
         mean_stress_family = mean(stressed[Wfamily==1], na.rm=TRUE),
         mean_stress_peers = mean(stressed[Wpeers==1], na.rm=TRUE),
         baseline_mean_stress_alone = mean(stressed[personal_time < 8 & alone ==1], na.rm=TRUE),
         baseline_mean_stress_family = mean(stressed[personal_time < 8 & Wfamily==1], na.rm=TRUE),
         baseline_mean_stress_peers = mean(stressed[personal_time < 8 & Wpeers==1], na.rm=TRUE),
         baseline_sd_stress = sd(stressed[personal_time < 8], na.rm=TRUE),
         baseline_alone_y = sum(alone[personal_time < 8], na.rm=TRUE),
         baseline_peer_y = sum(Wpeers[personal_time < 8], na.rm=TRUE),
         baseline_alone_n = sum(alone[personal_time < 8] == 0,  na.rm=TRUE),
         baseline_peer_n = sum(Wpeers[personal_time < 8] == 0,  na.rm=TRUE),
         baseline_alone_pct = baseline_alone_y / (baseline_alone_y + baseline_alone_n),
         baseline_peer_pct = sum(baseline_peer_y) / (baseline_peer_y + baseline_peer_n),
         alone_pct = sum(alone) / n(),
         stress_buffer_alone = mean(stressed[alone==1]) - mean(stressed[alone==0]),
         stress_buffer_family= mean(stressed[alone==1]) - mean(stressed[Wfamily==1]),
         stress_buffer_peer = mean(stressed[alone==1]) - mean(stressed[Wpeers==0])) %>%
  mutate(
         mean_stress_z = scale(mean_stress),
         mean_na_z = scale(mean_na),
         mean_pa_z = scale(mean_pa),
         alone_pct_z = scale(alone_pct),
         baseline_alone_pct_z = scale(baseline_alone_pct),
         baseline_peer_pct_z = scale(baseline_peer_pct),
         baseline_mean_stress_z = scale(baseline_mean_stress),
         baseline_mean_pa_z = scale(baseline_mean_pa),
         baseline_mean_na_z = scale(baseline_mean_na),
         baseline_mean_stress_alone_z = scale(baseline_mean_stress_alone),
         baseline_mean_stress_family_z = scale(baseline_mean_stress_family),
         baseline_mean_stress_peers_z = scale(baseline_mean_stress_peers),
         mean_stress_alone_z = scale(mean_stress_alone),
         mean_stress_family_z = scale(mean_stress_family),
         mean_stress_peers_z = scale(mean_stress_peers))

btw_summary_data = left_join(btw_summary_data, selfreport, by = 'ID')
btw_summary_data = mutate(btw_summary_data, stb = ifelse(Control == 0, 1, 0))

participants_events_first_7 = dplyr::filter(event_for_survival_analysis, survival_time_event < 8)
btw_summary_data = mutate(btw_summary_data, `Suicide Events\nWithin 6-Month Period`=ifelse(participant_suicide_event1==0, 'No Events', '1+ Events'),
                          suicide_event_for_jitter=ifelse(participant_suicide_event1==0, 0.01, 0.98))

btw_summary_data_ema_events = dplyr::filter(btw_summary_data, ! ID %in% participants_events_first_7$ID)
table(btw_summary_data_ema_events$participant_suicide_event1)
```

## Using EMA data *preceding* events (EMA data only from days 0-7, events days 8+)

```{r}
group_predict_events = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ stb + SITE,
                               family = binomial(link = 'logit'))

table(btw_summary_data_ema_events$participant_suicide_event1, btw_summary_data_ema_events$stb)

tab_model(group_predict_events)

ema_stress_predict_events = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_mean_stress_z + GROUP + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

ema_alone_pct_predict_events = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_alone_pct_z + GROUP + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

ema_peer_pct_predict_events = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_peer_pct_z + GROUP + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

ema_mean_na_predict_events = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_mean_na_z + GROUP + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

ema_mean_pa_predict_events = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_mean_pa_z + GROUP + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

ema_stress_predict_events_si = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_mean_stress_z + SSI21_i + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

ema_alone_pct_predict_events_si = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_alone_pct_z + SSI21_i + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

ema_mean_na_predict_events_si = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~baseline_mean_na_z + SSI21_i + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

ema_mean_pa_predict_events_si = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_mean_pa_z + SSI21_i + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

tab_model(stress_buffer_glm_level2, ema_stress_predict_events, ema_alone_pct_predict_events, ema_mean_na_predict_events, ema_mean_pa_predict_events,
          ema_stress_predict_events_si, ema_alone_pct_predict_events_si, ema_mean_na_predict_events_si, ema_mean_pa_predict_events_si)


ema_stress_coef = broom::tidy(ema_stress_predict_events, exponentiate=TRUE, conf.int=TRUE)
ema_na_coef = broom::tidy(ema_mean_na_predict_events, exponentiate=TRUE, conf.int=TRUE)
ema_pa_coef = broom::tidy(ema_mean_pa_predict_events, exponentiate=TRUE, conf.int=TRUE)
ema_alone_coef = broom::tidy(ema_alone_pct_predict_events, exponentiate=TRUE, conf.int=TRUE)
```

## Emmeans plotting
```{r}
stress_emmeans = emmeans::emmeans(ema_stress_predict_events, specs = 'baseline_mean_stress_z', 
                                  at = list(baseline_mean_stress_z = seq(from = min(btw_summary_data_ema_events$baseline_mean_stress_z, na.rm = TRUE), 
                                                                         to = max(btw_summary_data_ema_events$baseline_mean_stress_z, na.rm = TRUE), by = 0.05)),
                                  type = 'response') %>%
  as.data.frame()

stress_emmean_plt = ggplot(stress_emmeans, aes(x=baseline_mean_stress_z, y = prob)) +
  geom_line(lwd = 2) + 
  geom_jitter(data = btw_summary_data_ema_events, aes(x = baseline_mean_stress_z, y = suicide_event_for_jitter, 
                                                      color = `Suicide Events\nWithin 6-Month Period`), alpha = 0.3, height = 0.005, width = 0, size = 1) +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.2) +
  labs(y = 'P(Suicide Event)\nOver 6-Months', x = 'Momentary Stress\nFirst 7 Days', title = '') +
  theme_bw() +
  theme(text = element_text(face='bold')) +
  scale_color_manual(values = c('red', 'navy')) +
  theme(legend.position = 'none') +
  geom_text(x = 0, y = 0.75, label = paste0(
      'aOR= ', round(ema_stress_coef$estimate[ema_stress_coef$term=='baseline_mean_stress_z'], 2),
      ' 95% CI [', round(ema_stress_coef$conf.low[ema_stress_coef$term=='baseline_mean_stress_z'], 2),
      ', ',  round(ema_stress_coef$conf.high[ema_stress_coef$term=='baseline_mean_stress_z'], 2), 
      '] \np=', round(ema_stress_coef$p.value[ema_stress_coef$term=='baseline_mean_stress_z'], 3)
      ))  +
  ylim(0,1)

pa_emmeans = emmeans::emmeans(ema_mean_pa_predict_events, specs = 'baseline_mean_pa_z', 
                              at = list(baseline_mean_pa_z = seq(from = min(btw_summary_data_ema_events$baseline_mean_pa_z, na.rm = TRUE), 
                                                                 to = max(btw_summary_data_ema_events$baseline_mean_pa_z, na.rm = TRUE), by = 0.05)), 
                              type = 'response') %>%
  as.data.frame()


pa_emmean_plt = ggplot(pa_emmeans, aes(x=baseline_mean_pa_z, y = prob)) +
  geom_line(lwd = 2) + 
  geom_jitter(data = btw_summary_data_ema_events, aes(x = baseline_mean_pa_z, y = suicide_event_for_jitter, 
                                                      color = `Suicide Events\nWithin 6-Month Period`), alpha = 0.3, height = 0.005, width = 0, size = 1) +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.2) +
  labs(y = NULL, x = 'Positive Affect\nFirst 7 Days', title = '') +
  theme_bw() +
  theme(text = element_text(face='bold')) +
  scale_color_manual(values = c('red', 'navy')) +
  theme(legend.position = 'none') +
  geom_text(x = 0, y = 0.75, label = paste0(
      'aOR= ', round(ema_pa_coef$estimate[ema_pa_coef$term=='baseline_mean_pa_z'], 2),
      ' 95% CI [', round(ema_pa_coef$conf.low[ema_pa_coef$term=='baseline_mean_pa_z'], 2),
      ', ',  round(ema_pa_coef$conf.high[ema_pa_coef$term=='baseline_mean_pa_z'], 2), 
      '] \np=', round(ema_pa_coef$p.value[ema_pa_coef$term=='baseline_mean_pa_z'], 3)
      ))  +
  ylim(0,1)


na_emmeans = emmeans::emmeans(ema_mean_na_predict_events, specs = 'baseline_mean_na_z', 
                              at = list(baseline_mean_na_z = seq(from = min(btw_summary_data_ema_events$baseline_mean_na_z, na.rm = TRUE),
                                                                 to = max(btw_summary_data_ema_events$baseline_mean_na_z, na.rm = TRUE), by = 0.05)), 
                              type = 'response') %>%
  as.data.frame()


na_emmean_plt = ggplot(na_emmeans, aes(x=baseline_mean_na_z, y = prob)) +
  geom_line(lwd = 2) + 
  geom_jitter(data = btw_summary_data_ema_events, aes(x = baseline_mean_na_z, y = suicide_event_for_jitter, 
                                                      color = `Suicide Events\nWithin 6-Month Period`), alpha = 0.3, height = 0.005, width = 0, size = 1) +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.2) +
  labs(y = NULL, x = 'Negative Affect\nFirst 7 Days', title = '') +
  theme_bw() +
  theme(text = element_text(face='bold')) +
  scale_color_manual(values = c('red', 'navy')) +
  theme(legend.position = 'none') +
  geom_text(x = 0.1, y = 0.75, label = paste0(
      'aOR= ', round(ema_na_coef$estimate[ema_na_coef$term=='baseline_mean_na_z'], 2),
      ' 95% CI [', round(ema_na_coef$conf.low[ema_na_coef$term=='baseline_mean_na_z'], 2),
      ', ',  round(ema_na_coef$conf.high[ema_na_coef$term=='baseline_mean_na_z'], 2), 
      '] \np=', round(ema_na_coef$p.value[ema_na_coef$term=='baseline_mean_na_z'], 3)
      ))  +
  ylim(0,1)

alone_pct_emmeans = emmeans::emmeans(ema_alone_pct_predict_events, specs = 'baseline_alone_pct_z', 
                                     at = list(baseline_alone_pct_z = seq(from = min(btw_summary_data_ema_events$baseline_alone_pct_z, na.rm = TRUE), 
                                                                          to = max(btw_summary_data_ema_events$baseline_alone_pct_z, na.rm = TRUE), by = 0.05)), 
                                     type = 'response') %>%
  as.data.frame()

 
alone_pct_emmean_plt = ggplot(alone_pct_emmeans, aes(x=baseline_alone_pct_z, y = prob)) +
  geom_line(lwd = 2) + 
  geom_jitter(data = btw_summary_data_ema_events, aes(x = baseline_alone_pct_z, y = suicide_event_for_jitter, 
                                                      color = `Suicide Events\nWithin 6-Month Period`), alpha = 0.3, height = 0.005, width = 0, size = 1) +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.2) +
  labs(y = NULL, x = '% of Prompts Spent Alone\nFirst 7 Days', title = '') +
  theme_bw() +
  theme(text = element_text(face='bold')) +
  scale_color_manual(values = c('red', 'navy')) +
  theme(legend.position = 'none') +
  geom_text(x = 0.5, y = 0.75, label = paste0(
      'aOR= ', round(ema_alone_coef$estimate[ema_alone_coef$term=='baseline_alone_pct_z'], 2),
      ' 95% CI [', round(ema_alone_coef$conf.low[ema_alone_coef$term=='baseline_alone_pct_z'], 2),
      ', ',  round(ema_alone_coef$conf.high[ema_alone_coef$term=='baseline_alone_pct_z'], 2), 
      '] \np=', round(ema_alone_coef$p.value[ema_alone_coef$term=='baseline_alone_pct_z'], 3)
      ))  +
  ylim(0,1)

ema_predict_suicide_events_grid = cowplot::plot_grid(stress_emmean_plt, alone_pct_emmean_plt, na_emmean_plt, pa_emmean_plt, 
                                                     nrow = 1, rel_widths= c(1.1, 1, 1, 1),
                                                     labels = c('A', 'B', 'C', 'D'))


cowplot::save_plot(ema_predict_suicide_events_grid, filename = '../figures/main/3_ema_predict_suicide_events.png', base_height = 5, base_width = 12)
```


## Models only among participants with stb history
```{r}
ema_stress_predict_events_stb = glm(data = btw_summary_data_ema_events %>% dplyr::filter(stb==1), participant_suicide_event1 ~ baseline_mean_stress_z + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

ema_alone_pct_predict_events_stb = glm(data = btw_summary_data_ema_events%>% dplyr::filter(stb==1), participant_suicide_event1 ~ baseline_alone_pct_z + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

ema_mean_na_predict_events_stb = glm(data = btw_summary_data_ema_events%>% dplyr::filter(stb==1), participant_suicide_event1 ~ baseline_mean_na_z + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

ema_mean_pa_predict_events_stb = glm(data = btw_summary_data_ema_events%>% dplyr::filter(stb==1), participant_suicide_event1 ~ baseline_mean_pa_z + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

tab_model(ema_stress_predict_events_stb, ema_alone_pct_predict_events_stb, ema_mean_na_predict_events_stb, ema_mean_pa_predict_events_stb)
```

# Covarying for strain
```{r}
ema_stress_predict_events_strain = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_mean_stress_z + GROUP + sex + SITE + age + iPhone + StressTH,
                               family = binomial(link = 'logit'))

ema_alone_pct_predict_events_strain = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_alone_pct_z + GROUP + sex + SITE + age + iPhone + StressTH,
                               family = binomial(link = 'logit'))

ema_peer_pct_predict_events_strain = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_peer_pct_z + GROUP + sex + SITE + age + iPhone + StressTH,
                               family = binomial(link = 'logit'))

ema_mean_na_predict_events_strain = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_mean_na_z + GROUP + sex + SITE + age + iPhone + StressTH,
                               family = binomial(link = 'logit'))

ema_mean_pa_predict_events_strain = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_mean_pa_z + GROUP + sex + SITE + age + iPhone + StressTH,
                               family = binomial(link = 'logit'))

tab_model(ema_stress_predict_events_strain, ema_alone_pct_predict_events_strain, ema_mean_na_predict_events_strain, ema_mean_pa_predict_events_strain)

```


# Prediction of suicidal ideation
```{r}
ema_stress_predict_si= lm(data = btw_summary_data_ema_events, scale(SSI19_6) ~ baseline_mean_stress_z + GROUP + sex + SITE + age + iPhone)

ema_alone_pct_predict_si = lm(data = btw_summary_data_ema_events, scale(SSI19_6) ~ baseline_alone_pct_z + GROUP + sex + SITE + age + iPhone)

ema_mean_na_predict_si = lm(data = btw_summary_data_ema_events, scale(SSI19_6) ~ baseline_mean_na_z + GROUP + sex + SITE + age + iPhone)

ema_mean_pa_predict_si = lm(data = btw_summary_data_ema_events, scale(SSI19_6) ~ baseline_mean_pa_z + GROUP + sex + SITE + age + iPhone)

tab_model(ema_stress_predict_si, ema_alone_pct_predict_si, ema_mean_na_predict_si, ema_mean_pa_predict_si)

```


# Prediction of depression
```{r}
ema_stress_predictmfq= lm(data = btw_summary_data_ema_events, scale(MFQ_6) ~ baseline_mean_stress_z + GROUP + sex + SITE + age + iPhone)

ema_alone_pct_predictmfq = lm(data = btw_summary_data_ema_events, scale(MFQ_6) ~ baseline_alone_pct_z + GROUP + sex + SITE + age + iPhone)

ema_mean_na_predictmfq = lm(data = btw_summary_data_ema_events,scale(MFQ_6) ~ baseline_mean_na_z + GROUP + sex + SITE + age + iPhone)

ema_mean_pa_predictmfq = lm(data = btw_summary_data_ema_events, scale(MFQ_6) ~ baseline_mean_pa_z + GROUP + sex + SITE + age + iPhone)

tab_model(ema_stress_predictmfq, ema_alone_pct_predictmfq, ema_mean_na_predictmfq, ema_mean_pa_predictmfq)

```



# Strain prediction of events

```{r}

strain_characteristics = c('CIAllCT', 'CIAllTH', 'CDAllCT', 'CDAllTH', 'CHAllCT', 'CHAllTH', 'CEAllCT', 'CEAllTH', 'CRAllCT',
                           'CRAllTH', 'StressCT', 'StressTH', 'EvntCT', 'DiffCT', 'EvntTH','DiffTH')


btw_summary_data = mutate(btw_summary_data,
                       # Count
                       RecDAllCT = RecDDiffCT + RecEvntCT,
                       RecEAllCT = RecEDiffCT + RecEvntCT,
                       RecHAllCT = RecHDiffCT + RecEvntCT,
                       RecLAllCT = RecLDiffCT + RecEvntCT,
                       RecRAllCT = RecRDiffCT + RecEvntCT,
                       # Severity
                       RecDAllTH = RecDDiffTH + RecEvntTH,
                       RecEAllTH = RecEDiffTH + RecEvntTH,
                       RecHAllTH = RecHDiffTH + RecEvntTH,
                       RecLAllTH = RecLDiffTH + RecEvntTH,
                       RecRAllTH = RecRDiffTH + RecEvntTH
                       )

strain_characteristics_pastyear = c('RecDAllCT', 'RecEAllCT', 'RecHAllCT', 'RecLAllCT', 'RecRAllCT',
                                    'RecDAllTH', 'RecEAllTH', 'RecHAllTH', 'RecLAllTH', 'RecRAllTH', 
                                    'RecTotCT', 'RecTotTH', 'RecEvntCT', 'RecEvntTH', 'RecDiffCT', 'RecDiffTH')


strain_event_models1 = btw_summary_data %>%
    pivot_longer(all_of(c(strain_characteristics, strain_characteristics_pastyear))) %>%
    group_by(name) %>%
    mutate(value_scale = scale(value)) %>%
    nest() %>%
    mutate(m = purrr::map(data, ~glm(data = ., family = binomial(link = 'logit'),
                                                       participant_suicide_event1 ~ value_scale + age + sex + SITE + iPhone)),
           model_coef= purrr::map(m, ~broom::tidy(., conf.int=TRUE, exponentiate=TRUE)),
           ctrl_baseline_si = 'No')

strain_event_models2 = btw_summary_data %>%
    pivot_longer(all_of(c(strain_characteristics, strain_characteristics_pastyear))) %>%
    group_by(name) %>%
    mutate(value_scale = scale(value)) %>%
    nest() %>%
    mutate(m = purrr::map(data, ~glm(data = ., family = binomial(link = 'logit'),
                                                       participant_suicide_event1 ~ value_scale + age + Control + sex + SITE + iPhone)),
           model_coef= purrr::map(m, ~broom::tidy(., conf.int=TRUE, exponentiate=TRUE)),
           ctrl_baseline_si = 'Yes')

strain_event_models = rbind(strain_event_models1, strain_event_models2)


strain_coef = strain_event_models %>%
  unnest(model_coef) %>%
  dplyr::select(-data, -m) %>%
  dplyr::filter(term == 'value_scale') %>%
  dplyr::mutate(characteristic = case_when(
        startsWith(name, 'CI') ~ 'Interpersonal Loss',
        startsWith(name, 'CD') ~ 'Physical Danger',
        startsWith(name, 'CH') ~ 'Humiliation',
        startsWith(name, 'CE') ~ 'Entrapment',
        startsWith(name, 'CR') ~ 'Role Change/Reversal',
        startsWith(name, 'Stress') ~ 'Total Stressors',
        startsWith(name, 'Evnt') ~ 'Total Acute Life Events',
        startsWith(name, 'Diff') ~ 'Total Chronic Difficulties',
        name == 'RecDAllTH' ~ 'Physical Danger',
        name == 'RecEAllTH' ~ 'Entrapment',
        name == 'RecHAllTH'~ 'Humiliation', 
        name == 'RecLAllTH' ~ 'Interpersonal Loss',
        name =='RecRAllTH' ~ 'Role Change/Reversal',
        name =='RecTotTH' ~ 'Total Stressors',
        name =='RecEvntTH' ~ 'Total Acute Life Events',
        name == 'RecDiffTH'~ 'Total Chronic Difficulties'
    ),
    strain_timing = ifelse(startsWith(name, 'Rec'), 'Past Year', 'Lifetime'),
    strain_type = case_when(
        endsWith(name, 'TH') ~ 'Severity',
         endsWith(name, 'CT') ~ 'Count'),
    strain_total = case_when(
      grepl('Total', characteristic) ~ 'Total Summaries',
      !grepl('Total', characteristic) ~ 'Core Social-Psych\nCharacteristics'
    )) %>%
  dplyr::filter(strain_type == 'Severity')


strain_events_plot = strain_coef %>%
  dplyr::filter(ctrl_baseline_si == 'Yes') %>%
  ggplot(data = , aes(x = characteristic, color = strain_timing, y = estimate)) +
  geom_point(position = position_dodge(0.3), size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0, 
                position = position_dodge(0.3),
                lwd = 1) +
  theme_bw() +
  coord_flip() +
  facet_grid(rows = vars(strain_total), scales = 'free_y') +
  geom_hline(yintercept = 1, lty = 2) +
  labs(color = 'Metric', x = NULL, y = 'Ajusted Odds Ratio',
       title = 'Associations between prior stress & adversity severity\n& suicide events within study period') +
  scale_color_manual(values = c('black', 'red4'))

ggsave(strain_events_plot, file = '../figures/supplement/strain_events_plot.png', height = 7, width = 7)
```

## Survival analysis


```{r}
event_for_survival_analysis = left_join(event_for_survival_analysis, btw_summary_data, by = 'ID')
event_for_survival_analysis = mutate(event_for_survival_analysis, baseline_mean_stress_z_split = baseline_mean_stress_z > median(baseline_mean_stress_z, na.rm=TRUE))

#event_for_survival_analysis = dplyr::mutate(event_for_survival_analysis, survival_time_event = ifelse(survival_time_event >250, 250, survival_time_event))

df_for_cox_models = dplyr::select(event_for_survival_analysis,ID, survival_time_event, 
                                  suicide_event1_participant, baseline_mean_stress_z_split, contains('baseline'), GROUP, SSI19_i, age, sex, SITE)

# TEST: group
gen_fit = survival::survfit(Surv(survival_time_event, suicide_event1_participant) ~ GROUP, data=df_for_cox_models)
autoplot(gen_fit)

# stress
stress_cox = coxph(Surv(survival_time_event, suicide_event1_participant) ~ baseline_mean_stress_z + GROUP + age + sex + SITE, data=df_for_cox_models)
stress_cox2 = coxph(Surv(survival_time_event, suicide_event1_participant) ~ baseline_mean_stress_z + SSI19_i + age + sex + SITE, data=df_for_cox_models)
stress_cox_medsplit = survival::survfit(Surv(survival_time_event, suicide_event1_participant) ~ baseline_mean_stress_z_split, data=df_for_cox_models)
stress_cox_medsplit_m = coxph(Surv(survival_time_event, suicide_event1_participant) ~ baseline_mean_stress_z_split + SSI19_i + age + sex + SITE, data=df_for_cox_models)



# PA
pa_cox = coxph(Surv(survival_time_event, suicide_event1_participant) ~ baseline_mean_pa_z + GROUP + age + sex + SITE, data=df_for_cox_models)
pa_cox2 = coxph(Surv(survival_time_event, suicide_event1_participant) ~ baseline_mean_pa_z + SSI19_i + age + sex + SITE, data=df_for_cox_models)

# NA
na_cox = coxph(Surv(survival_time_event, suicide_event1_participant) ~ baseline_mean_na_z + GROUP + age + sex + SITE, data=df_for_cox_models)
na_cox2 = coxph(Surv(survival_time_event, suicide_event1_participant) ~ baseline_mean_na_z + SSI19_i + age + sex + SITE, data=df_for_cox_models)

# alone
alone_pct_cox = coxph(Surv(survival_time_event, suicide_event1_participant) ~ baseline_alone_pct_z + GROUP + age + sex + SITE, data=df_for_cox_models)
alone_pct_cox2 = coxph(Surv(survival_time_event, suicide_event1_participant) ~ baseline_alone_pct_z + SSI19_i+ age + sex + SITE, data=df_for_cox_models)



tab_model(stress_cox, pa_cox, na_cox, alone_pct_cox, stress_cox_medsplit_m)
tab_model(stress_cox2, pa_cox2, na_cox2, alone_pct_cox2)


cox.zph(stress_cox, transform = 'km', terms = TRUE, singledf=FALSE, global=TRUE)

```

# Zero-inflated poisson regression predicting # of events

```{r}
btw_summary_data_ema_events = left_join(btw_summary_data_ema_events, dplyr::select(sum_suicide_events_participant, ID=ptid, n_events = n), by = 'ID')

btw_summary_data_ema_events = mutate(btw_summary_data_ema_events, n_events = ifelse(is.na(n_events), 0, n_events))


ggplot(btw_summary_data_ema_events, aes(x = baseline_mean_stress_z, y = n_events)) +
  geom_point() +
  geom_smooth(method='lm')


length_on_study = event_for_survival_analysis %>%
  mutate(length_in_study = as.numeric(end_date - start_date)) %>%
  dplyr::select(ID, length_in_study)

btw_summary_data_ema_events = left_join(btw_summary_data_ema_events, length_on_study, by = 'ID')


zip_stress1 = pscl::zeroinfl(n_events ~ baseline_mean_stress_z  +GROUP  + offset(log(length_in_study)) | baseline_mean_stress_z +GROUP+ offset(log(length_in_study)), data = btw_summary_data_ema_events)

summary(zip_stress1)
```

# Does positive affect buffer negative affect / stress? (interaction models)
```{r}
ema_stress_buffer = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_mean_stress_z*baseline_mean_pa_z + GROUP + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))

ema_na_buffer = glm(data = btw_summary_data_ema_events, participant_suicide_event1 ~ baseline_mean_na_z*baseline_mean_pa_z + GROUP + sex + SITE + age + iPhone,
                               family = binomial(link = 'logit'))


tab_model(ema_stress_buffer, ema_na_buffer, show.est = TRUE, transform = NULL)
```